Search-Based Test Data Generation for JavaScript Functions with DOM

### 2018-01-15 doc/icse2018/icse2018.tex
Another review round.

\usepackage{syntax}
\usepackage{multirow}
\usepackage{booktabs} % For formal tables
\usepackage{tikz}
\usetikzlibrary{arrows,shadows,positioning,decorations.pathmorphing}
\usepackage{pgf-umlsd}
\usepackage{tikz-uml}
\usepackage{hyperref}
\usepackage{upquote}
\usepackage{listings}
\usepackage{lsthaskell}
\usepackage{xspace}
\usepackage{algorithmic}
\usepackage{longtable}
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepackage[titlenotnumbered, linesnumbered, ruled]{algorithm2e} % language for algorithm description
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = red, %Colour of internal links
  citecolor    = green %Colour of citations
}

@@

\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings.\footnote{This is an abstract footnote}
\end{abstract}

@@

According to the Stack Overflow survey\footnote{\url{https://insights.stackoverflow.com/survey/2017}} run in 2017 across developers, JavaScript has been recognized for the firth time in row as the most commonly used programming language with Node.js as the most commonly used technology. Such a tremendous ever increasing success of JS happens thanks to various language features including dynamic typing, native browser support (DOM).

Say about popularity of JS applications including nodejs. Cite some works referring to challenges in testing.

Common mistakes JS developers do\footnote{\url{https://www.toptal.com/javascript/10-most-common-javascript-mistakes}}.

Notable errors in 2017\footnote{\url{https://www.fastcompany.com/40505226/the-year-that-software-bugs-ate-the-world}}.

List current practices in testing and show where they fail. Make connection with DOM.

The state of JS 2017 survey~\footnote{\url{https://stateofjs.com/}} contains an extensive list of testing frameworks including Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma and QUnit among the most used ones in the decreasing order of popularity. These frameworks support testing at different levels unit, integration and functional. The same survey also indicated relatively low satisfaction rate of the current state of JS testing tools, only 3.2 out of 5. As result, new testing frameworks such as Cypress and Storybook are constantly appearing to improve the situation. But despite of this effort, test data generation stays a missing component not being properly addressed by any of the widely used testing frameworks. Research community has actively been working towards making progress in this direction introducing such tools as Artemis~\cite{artemis2011}, JSConTest~\cite{heidegger2010contract}, Crawljax~\cite{mesbah2012crawling}, JSeft~\cite{mirshokraie2015jseft} and Jalangi~\cite{sen2013jalangi}.

Recent study of client-side JS bugs~\cite{ocariza2017study} has shown that majority (68\%) of JS faults are DOM-related and occur in the JS code itself. This fact indicates that a better

Introduce search-based test generation.

Recent empirical study of client-side JavaScript bugs~\cite{frolin:TSE16} has shown that 68\% of faults are DOM related.

@@

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function isGameFinished() {
  var obj = document.getElementById('sudoku'); |c \label{isGameFinished.getSudoku} |c
  var subDivs = obj.getElementsByTagName('DIV'); |c \label{isGameFinished.getDivs} |c
  var allOk = true;
  for (var no = 0; no < subDivs.length; no++) { |c \label{isGameFinished.inFor.begin} |c
    if (subDivs[no].className.indexOf('sudokuSquare') >= 0 |c \label{isGameFinished.if1.begin} |c 
        && !subDivs[no].style.backgroundColor) { 
      var spans = subDivs[no].getElementsByTagName('SPAN');
      if (spans[0].innerHTML != spans[1].innerHTML) { |c \label{isGameFinished.if2.begin} |c
        allOk = false; |c \label{isGameFinished.unfinished} |c
        break;
      } |c \label{isGameFinished.if2.end} |c
    } |c \label{isGameFinished.if1.end} |c
  } |c \label{isGameFinished.inFor.end} |c
  if (allOk) { |c \label{isGameFinished.if3.begin} |c 
    console.log('Congratulations! You did it'); |c \label{isGameFinished.finished} |c
  } |c \label{isGameFinished.if3.end} |c
}
\end{lstlisting}
  \caption{JS function \texttt{isGameFinished.js} from sudoku}
  \label{code.newGame}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
function isGameFinished() {
  trace.push(2);
  var obj = document.getElementById("sudoku");
  trace.push(3);
  var subDivs = obj.getElementsByTagName("DIV");
  trace.push(4);
  var allOk = true;
  loopMap[5] = function () {
    var no = 0;
    if (subDivs.length || subDivs.length == 0) {
      return Math.abs(no - subDivs.length);
    }
    return 1;
  }();
  for (var no = 0; no < subDivs.length; no++) {
    trace.push(5);
    trace.push(6);
    trace.push(7);
    if (subDivs[no].className.indexOf("sudokuSquare") >= 0 &&
        !subDivs[no].style.backgroundColor) {
      trace.push(8);
      branchDistance.push({ 
        label: 7, 
        distance: Math.min(
          abs(subDivs[no].className.indexOf("sudokuSquare"), 0) + _K_,
          absZero(subDivs[no].style.backgroundColor))
      });
      trace.push(9);
      var spans = subDivs[no].getElementsByTagName("SPAN");
      trace.push(10);
      if (spans[0].innerHTML != spans[1].innerHTML) {
      trace.push(11);
      branchDistance.push({
        label: 10,
        distance: abs(spans[0].innerHTML, spans[1].innerHTML)
      });
      trace.push(12);
      allOk = false;
      break;
      } else {
        trace.push(14);
        branchDistance.push({label: 10,distance: _K_});
      }
    } else {
      trace.push(15);
      branchDistance.push({
        label: 7,
        distance: abs(subDivs[no].className.indexOf("sudokuSquare"), 0) +
                  absNegZero(subDivs[no].style.backgroundColor)
      });
    }
  }
  trace.push(5);
  trace.push(16);
  if (allOk) {
    trace.push(17);
    branchDistance.push({label: 16, distance: absNegZero(allOk)});
    trace.push(18);
    console.log("Congratulations! You did it");
  } else {
    trace.push(19);
    branchDistance.push({label: 16 ,distance: absZero(allOk)});
  }
  trace.push(-1);
}
\end{lstlisting}
  \caption{Instrumented version of the function \texttt{isGameFinished.js} }
  \label{code.isGameFinished.instr}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<!-- T1 covers: (6,15), (16,17) -->
<html>
  <body>
    <div id='sudoku'></div>
</body>
</html>

<!-- T2 covers: (6,7), (7,14), (16,17) -->
<html>
  <body>
    <div id='sudoku'>
      <div></div>
    </div>
  </body>
</html>

<!-- T3 covers: (6,7), (7,9), (10,13), (16,17) -->
<html>
  <body>
    <div id='sudoku'>
      <div class='sudokuSquare'>
        <span></span>
        <span></span>
      </div>
    </div>
  </body>
</html>

<!-- T4 covers: (6,7), (7,9), (10,11), (16,18) -->
<html>
  <body>
    <div id='sudoku'>
      <div class='sudokuSquare'>
        <span></span>
        <span>TEST</span>
      </div>
    </div>
  </body>
</html>
  \end{lstlisting}
  \caption{Input DOM arguments for \texttt{isGameFinished.js}}
  \label{fig.isGameFinished.tests}
\end{figure}

Let us consider JS function \texttt{isGameFinished} in Figure~\ref{code.newGame} taken from the web application game \emph{sudoku}~\ref{ref.sudoku}. This function checks if a sudoku solution entered by the user is valid. First, it finds the game field element (line~\ref{isGameFinished.getSudoku}), retrieves all child divs (line~\ref{isGameFinished.getDivs}), and iterates over that collection. Each div element which corresponds to a sudoku input square (line~\ref{isGameFinished.if1.begin}) consists of two span elements. One contains a user input, and another one, which is hidden at the moment, stores the expected value for the sudoku square. Once, it finds a square where two values are not equal (line~\ref{isGameFinished.if2.begin}) to each other, the game is considered unfinished (line~\ref{isGameFinished.unfinished}) and finished otherwise (line~\ref{isGameFinished.finished}).

Suppose now that we would like to (unit) test this function in isolation to the rest of the application source code. We aim to maximize branch coverage as a common testing criteria~\cite{why.this.common}. That is for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we have to construct a test input which passes the branch and normally exists the function. JS function has \emph{explicit} input arguments (none is our case), but it can also has \emph{implicit} arguments such as DOM state. As assume that DOM is the only global JS object that can be referred from the body of FUT. So in order to test our function we have to construct appropriate DOM objects as test inputs.

@@

Figure~\ref{fig.isGameFinished.tests} presents four tests that together provide full coverage of the FUT. Let us look at how to construct the test that covers branch $(10,11)$. Line~\ref{isGameFinished.getSudoku} expects DOM to have an element with the id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop on line~{isGameFinished.inFor.begin}, this element should have at least one child of the type $div$. The condition of the first $if$ (line~\ref{isGameFinished.if1.begin}) requires the div to be of the class \textquotesingle\texttt{sudokuSquare}\textquotesingle and have no background color. The last if-condition (line~\ref{isGameFinished.if2.begin}) expects that the div has two \emph{span} elements whose \emph{innerHtml} values are not equal. $T4$ in Figure~\ref{fig.isGameFinished.tests} shows a concrete example of DOM that meets all the conditions above.

As we have just seen even for such a relatively simple example, the construction of the test input is a far from trivial procedure because it requires deep understanding the program's semantics. In practice, web developers often have to deal with the gazillion of the JS frameworks and libraries freely available on the GitHub. This code is commonly untested and poorly documented. At the same time, unit testing is the most fundamental and easiest to provide level of testing. It greatly saves developers from the unexpected regression and creates a safety belt around the application.

@@

      \end{sdblock}
    \end{call}
  \end{sequencediagram}
  \caption{Architecture of the test framework}
  \label{fig.framework.architect} 
\end{figure}

Given JS function $f$ and a set of branches $\mathbb{B}_f$, the ultimate goal of the testing framework is to find input arguments for $f$ that cover all branches from $B_f$. We implemented such a framework as a client-server application which uses under the hood genetic test generation engine. The general workflow of the framework is depicted as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of the four interacting components \User, \Server, \Validator, and \Client. \Server is the key component of the whole framework. It is responsible for the \emph{Initialization Phase} and test data generation in the \emph{Genetic Phase}. Both phases will be covered in details in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively. Each newly constructed HTML document has to be processed by the \Validator\footnote{\url{https://validator.w3.org/nu/}} which checks its syntactical correctness and reports back the set of identified errors. \Server rejects the document if any errors were revealed and repeats the current step until the generated document has no errors anymore. \Client is a \emph{Node.js}\footnote{\url{https://nodejs.org/en/}} application that is mainly responsible for the execution of the JS code. We use the \emph{jsdom}\footnote{\url{https://github.com/tmpvar/jsdom}} library to model the virtual DOM and simulate native browser API calls.

Let us give a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. As an input to our testing framework, \User has to provide the JS file \emph{fun.js} that contains the function under test (FUT) \emph{fun} annotated with the type signature. Additionally, \User specifies a configuration file \emph{config} that defines various parameters for random data generation, genetic phase, logging etc. Experimentally, we have found an optimal set of input configurations that was used across all the experimental subjects in Section~\ref{sec.evaluation}. The received data is processed by \Server during the \emph{Initialization Phase}. At the end of it we get the following information back: the control flow graph (\emph{cfg}) of \emph{fun}, the list of its branches (\emph{branches}), the \emph{fun}'s  constant (\emph{cinfo}) info, the instrumented version \emph{ifun} of \emph{fun}, and the \emph{fun}'s signature (\emph{sig}). \Server sends a \texttt{POST} request to the \Client to store there \emph{ifun} and \emph{sig}. Depending on the chosen execution mode, the \Server either asks the \User to select the target branch or covers all branches one by one in a fixed order. For each branch our framework enters the \emph{Genetic Loop (Phase)}. Inside of this phase, the framework starts with a randomly generated population (\emph{Random Loop}), and continues to evolve the population (\emph{Fitness Loop}) until either a perfect entity is found or a certain termination criteria is achieved, e.g. the limit of generation attempts is reached. Each candidate in the population has to be evaluated in \emph{Evaluation Loop} by \Client. It matches the candidate $p$ according the known type signature \emph{sig} and invokes the instrumented JS function \emph{ifun(p)}. The execution trace along with some collected information are sent back to the \Server, where the final fitness value is computed. The new population is constructed by combining the results of the \emph{Crossover} and \emph{Mutation Loops}. If an HTML document is produced during one the genetic phases, it has to be approved by the \Validator first. At the end of the \emph{Fitness Loop} the candidate with the ``best'' fitness value is reported back to the \User. Once all branches are exercised, the algorithm reaches the end. In the follow up two sections we elaborate on the details of each individual component of our test generation framework.

@@

During the \emph{Initialization Phase} we analyze FUN and collect data necessary for the consecutive genetic testing phase. The main steps of the initialization are illustrated in Algorithm~\ref{alg.init}. Given a FUT \emph{fun.js} as an input, we parse the function and its type signature in line~\ref{alg.init.parse.fun.sig}. In order to maximize the chances of generating the fittest candidate in the genetic phase, we perform a static analysis of the FUT with the purpose to collect constant info in line~\ref{alg.init.collect.info}. Such constants include string and numeric literals as well as the arguments to the following DOM API methods:
\begin{itemize}
\item \texttt{getElementsByTagName}
\item \texttt{getElementById}
\item \texttt{getElementsByClassName}
\item \texttt{getElementsByName}
\end{itemize}
which respectively signify the reference to the \emph{tags}, \emph{ids}, \emph{classes} and \emph{names} used in the FUT. Next step is normalizing the FUT before applying any instrumentation (line~\ref{alg.init.transform.ast}), e.g. we need to transform \emph{if-then} branches into \emph{if-then-else}. We need to instrument the FUT (line~\ref{alg.init.instr}) in order be able to collect the run-time data such as the execution trace, the approach level, the branch and loop distances, and the constants. Upon the $ifun$ execution this information is gathered and fed back to the fitness evaluation procedure. The last two steps are to construct the control-flow graph (line~\ref{alg.init.build.cfg}) and identify all branches of the FUT (line~\ref{alg.init.get.branches}).

@@

The grammar of JS type annotations supported by our test generation framework is shown in Figure~\ref{fig.js.type.annot}. Each type annotation is a signature surrounded with a specialized JS type comment (\texttt{/*t...*/}). We distinguish two between kinds of signatures those who refer to the global DOM and ``ordinary'' JS type signatures. Functions with a DOM signature expect a \texttt{dom} argument being passed as the first implicit parameter to the function call. Ordinary JS signature is a non-zero sequence of JS types. The type can be \emph{Void}, \emph{Primitive} (\texttt{bool}, \texttt{int}, \texttt{float}, \texttt{string}) or \emph{Array} (recursive homogeneous array type).

@@

We want to generate HTML as it is specified by WHATWG\footnote{\url{https://html.spec.whatwg.org}}. HTML is an XML-family markup language that uses special tags and attributes to describe document structures on the Web. Recent HTML5 standard defines more than hundred of tags and attributes which both have to comply with extremely complex logic to constitute a syntactically valid HTML. Browsers are usually tolerant to syntactic errors in HTML. They will silently fix them and display the page anyway.

In order to support the variability of HTML content, we have developed an eDSL that prescribes the generation of syntactically valid HTML. This was a challenge on its own because, as we have mentioned early, the complexity of HTML and lack of formal specification. Our language is inspired by QuickCheck~\cite{claessen2011quickcheck} and essentially employs the library\footnote{\url{https://hackage.haskell.org/package/QuickCheck}} to define an \texttt{Arbitrary} instance for the \texttt{Html} data type. With some technical details left out for the sake of simplicity, Figure~\ref{fig.html.arb.def} highlights that definition. Data type \texttt{HtmlState} (lines~\ref{ref.HtmlState.begin}--\ref{ref.HtmlState.end}) describes the internal state of \texttt{Html} during generation. The state is composed of the following components:
\begin{itemize}
\item \texttt{depth} and \texttt{degree} specify maximum depth and node degree of HTML respectively
\item \texttt{ctx} is the current value of the HTML content model stack 
\item \texttt{tagFreq} provides statistical information about the frequency of tags in HTML
\item \texttt{tags}, \texttt{ids} and \texttt{classes} are predefined sets of HTML elements which are given preference in the generation process.  
\end{itemize}
Combined with the random generation monad \texttt{Gen}, this state is wrapped into a state monad transformer~\cite{jones1995functional} (line~\ref{ref.GenState}) which returns the randomly generated \texttt{Html} (line~\ref{ref.GenHtmlState}). Each HTML tag should be defined as a value of type \texttt{GenHtmlState}. Lines~\ref{ref.generateHTML.begin}--\ref{ref.generateHTML.end} show an example of content generation for the \texttt{<html>} tag. According to its specification\footnote{\url{https://html.spec.whatwg.org/multipage/semantics.html\#the-html-element}}, the tag is composed of the \texttt{<head>} tag followed by the \texttt{<body>} tag. First, we store the state of HTML generation and retrieve the current the depth (lines~\ref{ref.generateHTML.getState} and \ref{ref.generateHTML.getDepth} respectively). Then, if depth is equal to zero, we just return some string constant (line~\ref{ref.generateHTML.returnString}). Otherwise, we update the state with the depth decreased by one, generate new content for \texttt{head} and \texttt{body}, and return their combination (lines~\ref{ref.generateHTML.updateState}, \ref{ref.generateHTML.newHead}, \ref{ref.generateHTML.newBody} and \ref{ref.generateHTML.end} respectively).

@@

In essence, HTML document is a labeled tree with \emph{tags} instead of nodes and \emph{attributes} as labels. Therefore crossover and mutations are defined similar to those concepts for the expression trees in genetic programming.

\textbf{HTML Crossover} takes two HTML trees, randomly picks a node in each tree, and returns the first tree with a sub-tree swapped from the second tree.

>>>>

\usepackage{syntax}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{booktabs} % For formal tables
\usepackage{tikz}
\usetikzlibrary{arrows,shadows,positioning,decorations.pathmorphing}
\usepackage{pgf-umlsd}
\usepackage{tikz-uml}
\usepackage{hyperref}
\usepackage{upquote}
\usepackage{listings}
\usepackage{lsthaskell}
\usepackage{xspace}
\usepackage{algorithmic}
\usepackage{longtable}
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepackage[titlenotnumbered, linesnumbered, ruled]{algorithm2e} % language for algorithm description
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = red, %Colour of internal links
  citecolor    = green %Colour of citations
}

@@

\begin{abstract}
As of 2017 JavaScript language and Node.js platform continues to stay the most prevalent technologies voted by the developers in Stack Overflow survey. At the same time  
\end{abstract}

@@

According to the Stack Overflow survey\footnote{\url{https://insights.stackoverflow.com/survey/2017}} run in 2017 across developers, JavaScript has been recognized for the firth time in row as the most commonly used programming language with Node.js as the most commonly used technology. It is used through all technology stack web, back-ends and mobile. Such a tremendous ever increasing JS success is happening thanks to extensive dynamic features of the language, JS permissiveness and native browser support. On the contrary, JS applications are error-prone~\cite{why}, hard to analyze~\cite{why} and test~\cite{why}. Therefore, it is not surprising that year over year we hear new stories\footnote{\url{https://www.fastcompany.com/40505226/the-year-that-software-bugs-ate-the-world}} about another critical bug being found in a web application by Facebook, Twitter etc.

%% Common mistakes JS developers do\footnote{\url{https://www.toptal.com/javascript/10-most-common-javascript-mistakes}}.

%% Notable errors in 2017\footnote{\url{https://www.fastcompany.com/40505226/the-year-that-software-bugs-ate-the-world}}.

The state of JS 2017 survey~\footnote{\url{https://stateofjs.com/}} contains an extensive list of testing frameworks including Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma and QUnit among the most used ones in the decreasing order of popularity. These frameworks support testing at different levels unit, integration and functional. The same survey also indicated relatively low satisfaction rate of the current state of JS testing tools, only 3.2 out of 5. As result, new testing frameworks such as Cypress and Storybook are constantly appearing to improve the situation. But despite of this effort, test data generation stays a missing component not being properly addressed by any of the widely used testing frameworks. Research community has actively been working towards making progress in this direction introducing such tools as Artemis~\cite{artemis2011}, JSConTest~\cite{heidegger2010contract}, Crawljax~\cite{mesbah2012crawling}, JSeft~\cite{mirshokraie2015jseft} and Jalangi~\cite{sen2013jalangi}. Recent study of client-side JS bugs~\cite{ocariza2017study} has shown that majority (68\%) of JS faults are DOM-related and occur in JS code itself. This fact indicates that yet a better testing approach is needed when it comes to unit testing of JS.

Over the past two decades search-based test data generation has proven to be one of the most effective solutions for a wide range of technologies including object-oriented~\cite{paper}, web~\cite{paper} and mobile~\cite{paper}. In the recent years, some effort was done to apply search-based testing for dynamic languages such as Python~\cite{paper}, Ruby~\cite{paper} (maybe mention Go and Lua), but to the best of our knowledge search-based testing approach has never been implemented for JavaScript. This circumstance motivated us to introduce a search-based test generation framework for JS with presence of DOM. Our work makes the following contributions:
\begin{enumerate}
\item JS search-based testing framework
\item random HTML generation
\item genetic with restart algorithm
\item empirical evaluation 
\end{enumerate}

The paper is structured as follows:

@@

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function isGameFinished() {
  var obj = document.getElementById('sudoku'); |c \label{isGameFinished.getSudoku} |c
  var subDivs = obj.getElementsByTagName('DIV'); |c \label{isGameFinished.getDivs} |c
  var allOk = true;
  for (var no = 0; no < subDivs.length; no++) { |c \label{isGameFinished.inFor.begin} |c
    if (subDivs[no].className.indexOf('sudokuSquare') >= 0 |c \label{isGameFinished.if1.begin} |c 
        && !subDivs[no].style.backgroundColor) { 
      var spans = subDivs[no].getElementsByTagName('SPAN');
      if (spans[0].innerHTML != spans[1].innerHTML) { |c \label{isGameFinished.if2.begin} |c
        allOk = false; |c \label{isGameFinished.unfinished} |c
        break; 
      }
    }
  } |c \label{isGameFinished.outFor} |c
}
\end{lstlisting}
  \caption{JS function \texttt{isGameFinished.js} from sudoku}
  \label{code.isGameFinished}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<!-- T1: (5,5) -->            <!-- T2: (5,6), (7,15) -->
<html>                        <html>
 <body>                        <body>
  <div id='sudoku'>             <div id='sudoku'>
                                 <div></div>                           
  </div>                        </div> 
 </body>                       </body>    
</html>                       </html>

<!-- T3: (7,8), (10,14) -->   <!-- T4: (7,8), (10,11) -->
<html>                        <html>
 <body>                        <body>
  <div id='sudoku'>             <div id='sudoku'>
   <div class='sudokuSquare'>    <div class='sudokuSquare'>
    <span></span>                 <span></span> 
    <span></span>                 <span>TEST</span>
   </div>                        </div> 
  </div>                        </div>
 </body>                       </body>
</html>                       </html>
  \end{lstlisting}
  \caption{Input DOM arguments for \texttt{isGameFinished.js}}
  \label{fig.isGameFinished.tests}
\end{figure}

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from a web game \emph{sudoku}\footnote{\url{http://www.dhtmlgoodies.com/scripts/game_sudoku/game_sudoku.html}}. This function checks the validity of a sudoku solution given by a user. First, we find a DOM element representing the game field (line~\ref{isGameFinished.getSudoku}). Then, we collect all its child DIVs (line~\ref{isGameFinished.getDivs}). Each DIV element corresponds to a sudoku input square (line~\ref{isGameFinished.if1.begin}) and consists of two SPANs. First span contains a user input, whereas the second one, which is hidden by default, stores the expected value for that sudoku square. Inside of a for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate over the whole collection of input squares. Once, we find a square with two unequal SPAN values (line~\ref{isGameFinished.if2.begin}), the loop is terminated and the game is considered unfinished (line~\ref{isGameFinished.unfinished}) and finished otherwise.

Suppose now that we would like to unit test this function in isolation to the rest of the application. We aim to maximize the branch coverage as a common testing criteria~\cite{why.this.common}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. JS function takes \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as the DOM state. We assume that DOM is the only global variable that can be referenced in the body of the FUT. So in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments.

@@

Figure~\ref{fig.isGameFinished.tests} presents four tests that together deliver the full coverage of the FUT. Let us look now at how following the structure of the FUT in Figure~\ref{code.isGameFinished}, we can construct the $T4$ test covering the $(10,11)$ branch. Line~\ref{isGameFinished.getSudoku} expects the input DOM to have an element with the id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, that element should have at least one child \emph{DIV} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires this $div$ to be of the class \textquotesingle\texttt{sudokuSquare}\textquotesingle\ and also have no background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) expects that the $div$ consists of two \emph{SPAN} elements whose \emph{innerHtml} values are not equal. The resulted test $T4$ is shown in Figure~\ref{fig.isGameFinished.tests} and meets all the above conditions.

As we have just seen, even for such a relatively simple example, the construction procedure of a test input is far from trivial because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with the gazillion of the JS frameworks and libraries freely available on the GitHub whose code is commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and, at the same time, it is easiest to implement. It greatly saves the developers from unexpected regression and puts a safety belt around the application.

@@

      \end{sdblock}
    \end{call}
  \end{sequencediagram}
  \caption{The architecture of the test framework}
  \label{fig.framework.architect} 
\end{figure}

Given a JS function $f$ and a set of branches $\mathbb{B}_f$, the ultimate goal of the testing framework is to find input arguments for $f$ that cover all branches from $B_f$. We implemented such a framework as a client-server application which under the hood uses the genetic engine for test generation. The general workflow of the framework is depicted as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of the four interacting components \User, \Server, \Validator, and \Client. \Server is the key component of the whole framework. It is responsible for the \emph{Initialization Phase} and the \emph{Genetic Phase}. Both phases will be covered in details in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively.

Each newly constructed HTML document has to be processed by the \Validator\footnote{\url{https://validator.w3.org/nu/}} which checks its syntactic correctness and reports back the set of identified errors. \Server rejects the document if any errors were revealed and generates a new one until the generated document has no errors anymore. \Client is a \emph{Node.js}\footnote{\url{https://nodejs.org/en/}} application that is mainly responsible for the execution of the JS code. We use the \emph{jsdom}\footnote{\url{https://github.com/tmpvar/jsdom}} library to model the virtual DOM and simulate the native browser API calls.

Let us give a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. As an input for our testing framework, \User has to provide the JS file \emph{fun.js} that contains the function under test (FUT) \emph{fun} annotated with the type signature. Additionally, \User specifies a configuration file \emph{config} that defines various parameters for random data generation, genetic phase, logging etc. Experimentally, we have found an optimal set of input configurations that was used across all experimental objects in Section~\ref{sec.evaluation}. This data is processed by the \Server during the \emph{Initialization Phase}. As an output, we get back the following information about \emph{fun}: the control flow graph (\emph{cfg}), the complete list of its branches (\emph{branches}), the constant info (\emph{cinfo}), the instrumented version \emph{ifun}, and the signature (\emph{sig}). \Server sends a \texttt{POST} request with \emph{ifun} and \emph{sig} to the \Client. Depends on the configured execution mode, the \Server either asks the \User to select a \emph{target} branch or covers all branches one by one in a fixed order. For each branch, our framework enters the \emph{Genetic Loop (Phase)}. Inside of this phase, the framework starts with a randomly generated population (\emph{Random Loop}), and continues to evolve that population (\emph{Fitness Loop}) until either a perfect entity is found or a certain termination criteria is achieved, e.g. the limit on the number of generation attempts. Each candidate in the population has to be evaluated in \emph{Evaluation Loop} by \Client. It matches the candidate $p$ according to the type signature \emph{sig} and invokes the instrumented JS function on it: \emph{ifun(p)}. The execution trace along with some supplementary information about the execution are sent back to the \Server, where the final \emph{fitness value} is computed. The new population is constructed by combining the results of the \emph{Crossover} and \emph{Mutation Loops}. If a new HTML document is produced somewhere in the genetic phase, it has to be first approved by the \Validator. At the end of the \emph{Fitness Loop}, the candidate with the ``best'' fitness value is reported back to the \User. Once all the branches are exercised, the algorithm reaches the end. In the follow up two sections, we elaborate on the details of each individual phase of our test generation framework.

@@

During the \emph{Initialization Phase} we analyze the FUN in order to retrieve the data required for the consecutive genetic testing. The main steps of this phase are illustrated by Algorithm~\ref{alg.init}. Given a FUT \emph{fun.js} as an input, we parse the function and its type signature in line~\ref{alg.init.parse.fun.sig}. In order to maximize the chances of generating the fittest candidate in the genetic phase, in line~\ref{alg.init.collect.info} we perform static analysis of the FUT with the purpose to collect constant info. Such constants include the numeric and string literals, as well as the arguments to the following DOM API methods:
\begin{multicols}{2}
\footnotesize
\begin{itemize}
\item \texttt{getElementsByTagName}
\item \texttt{getElementById}
\item \texttt{getElementsByClassName}
\item \texttt{getElementsByName}
\end{itemize}
\end{multicols}
which respectively signify the reference to the \emph{tags}, \emph{ids}, \emph{classes} and \emph{names} encountered in the FUT. Next step is normalizing the FUT before applying any instrumentation (line~\ref{alg.init.transform.ast}), e.g. transform \emph{if-then} branches into \emph{if-then-else}. We instrument the FUT (line~\ref{alg.init.instr}) in order to collect the run-time data such as the execution trace, the approach level, the branch and loop distances, and the constants. Upon the $ifun$ execution this information is gathered and fed back to the fitness evaluation procedure. Finally, we construct the control-flow graph (line~\ref{alg.init.build.cfg}) and identify all branches of the FUT (line~\ref{alg.init.get.branches}).

@@

The grammar of the JS type annotations supported by our test generation framework is shown in Figure~\ref{fig.js.type.annot}. Each type annotation is a signature embedded into a specialized JS type comment (\texttt{/*t...*/}). We distinguish between two kinds of signatures those which refer to the global DOM and the ``ordinary'' JS type signatures. The functions of the former signature expect some \texttt{dom} argument to be passed as the first implicit parameter for the function call. Whereas ordinary JS signature is a non-zero sequence of the JS types. The type can be \emph{Void}, \emph{Primitive} (\texttt{bool}, \texttt{int}, \texttt{float}, \texttt{string}) or \emph{Array} (recursive homogeneous array type).

\subsubsection{Instrumentation}
\label{sub.sec.instrument}

Figure~\ref{code.isGameFinished.instr} presents the \texttt{isGameFinished} function after instrumentation. The global variable \emph{trace} in line~\ref{isGameFinished.instr.trace} is an array of labels corresponding to the sequence of executed program statements. All statements have to be traced because each one, in theory, can rise an exception. The \texttt{loopMap} variable in line~\ref{isGameFinished.instr.loopMap} approximates the potential number of iterations for the consecutive for-loop. If no exact bound can be computed, we set it to one. The \texttt{branchDistance} map tracks the distances to the deviated branches according to the translation formulae~\cite{tracey1998automated}.

TODO: maybe expand this section with the JS type casting for branch distance computation

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
function isGameFinished() {
 trace.push(2); |c \label{isGameFinished.instr.trace} |c
 var obj = document.getElementById("sudoku");
 trace.push(3);
 var subDivs = obj.getElementsByTagName("DIV");
 trace.push(4);
 var allOk = true;
 loopMap[5] = function () { |c \label{isGameFinished.instr.loopMap} |c
   var no = 0;
   if (subDivs.length || subDivs.length == 0) {
     return Math.abs(no - subDivs.length);
   }
   return 1;
 }();
 trace.push(5);
 for (var no = 0; no < subDivs.length; no++) {
  trace.push(6);
  trace.push(7);
  if (subDivs[no].className.indexOf("sudokuSquare") >= 0 &&
      !subDivs[no].style.backgroundColor) {
   trace.push(8);
   branchDistance.push({ 
    label: 7, 
    distance: Math.min(
     abs(subDivs[no].className.indexOf("sudokuSquare"), 0) + _K_, 
     absZero(subDivs[no].style.backgroundColor))
   });
   trace.push(9);
   var spans = subDivs[no].getElementsByTagName("SPAN");
   trace.push(10);
   if (spans[0].innerHTML != spans[1].innerHTML) {
    trace.push(11);
    branchDistance.push({
     label: 10,
     distance: abs(spans[0].innerHTML, spans[1].innerHTML)
    });
    trace.push(12);
    allOk = false;
    break;
   } else {
     trace.push(14);
     branchDistance.push({label: 10,distance: _K_});
   }
  } else {
   trace.push(15);
   branchDistance.push({
    label: 7,
    distance: 
     abs(subDivs[no].className.indexOf("sudokuSquare"), 0) + 
     absNegZero(subDivs[no].style.backgroundColor)
   });
  }
 }
 trace.push(5);
 trace.push(-1);
}
\end{lstlisting}
  \caption{Instrumented version of \texttt{isGameFinished.js} }
  \label{code.isGameFinished.instr}
\end{figure}

@@

@@

In essence, HTML document is a labeled tree with \emph{tags} instead of nodes and \emph{attributes} as labels. Therefore, the crossover and mutations are defined similar to those operators for the \emph{expression trees} in genetic programming.

\textbf{HTML Crossover} takes two HTML tree candidates, randomly picks a node in the first one, and replaces the respective sub-tree by a randomly selected sub-tree from the second candidate.
